### 2018-08-06 hitting_times.Rmd
hitting time theory

# Finding $G$ from $H$

Suppose that $G$ is the generator matrix of an irreducible continuous-time Markov chain $X$,
so that $G_{ij} \ge 0$ for $i \neq j$ and
$$ G 1 = 0 $$.

Let $\tau_{j} = \inf\{t \ge 0 : X_t = j\}$ and $H_{ij} = \E[\tau_j \,|\, X_0 = i]$.
Then we know that
$$
    (G H)_{ij} = -1 \qquad \text{for} \; i \neq j ,
$$
i.e.,
$$
    GH = - 1 1^T + \diag(x).
$$

What is $x$?  Well, note that
$$
    G = (\diag(x) - 1 1^T) H^{-1}
$$
and so
$$ \begin{aligned}
    0 &= G1 \\
    &= (\diag(x) - 1 1^T) H^{-1} x 
\end{aligned} $$
and so
$$
    x_i = \frac{ 1^T H^{-1} 1 }{ (H^{-1} 1)_i } .
$$
Furthermore,
the [Random Target Lemma](https://www.stat.berkeley.edu/users/aldous/RWG/Book_Ralph/Ch2.S2.html)
tells us that $H \pi = 1$, so thus $H^{-1} 1 = \pi$, the stationary distribution of the chain.
Therefore, $x_i = 1/\pi_i$,
and in fact $1^T H^{-1} 1 = 1$.

In summary,
$$
    G = (\diag(1/H^{-1} 1) - 1 1^T) H^{-1} .
$$

# Finding $G$ from $R$

The resistance distance is given by
$R = H + H^T$, so if we can modify $H$ by a skew-symmetric matrix,
we won't change $R$.
Can we do this and retain a valid $G$?
I.e., does $R$ uniquely specify $G$?
Let $Z + Z^T = 0$. 
Since $G$ is a continuous function of $H$, if all entries of $G$ are nonzero
we can find a small enough $\epsilon$ such that the $G'$ corresponding to $H' = H + \epsilon Z$
is also a valid generator matrix.
If $G$ has zero entries this may not be true.
To find a direction that we can change $H$ in that doesn't mess this up,
we could differentiate $dG/dH$.